Near instantaneous response of mesophyll conductance to light availability allows shade leaves to lie in wait

### 2015-08-03 manuscript/manuscript_wtc3.Rmd
more edits to results

##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
Leaf gas exchange measurements were performed monthly through to the final harvest.  Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.  The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy. In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.  As shade leaves most likely developed slower this assured that older leaves in the lower canopy were avoided. The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
\
Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.  A hand-held quantum sensor () was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade. A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type. Five ceptometer readings were recorded within the canopy at random locations of the height of each selected leaf.  The mean of these readings was assumed to represent the local light environment of full sun and shade leaves for each tree. All measurements of PPFD and gas exchange were performed on full sun days between 10:00-14:30 h.

@@

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite. Next, the observed discrimination ($\Delta$~obs~) is calculated from [@evans1986carbon]:

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air~(4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$  is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance **(von Caemmerer 1981). The ternary effect corrections (t) are desribed by:

@@

@@

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R with WTC as a random effect. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models of leaf photosynthesis as a functions of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R. For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R.  Results were considered significant at P&leq;0.05.

@@

## Leaf resource distribution
Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`). The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).  This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types. Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p < 0.001) by ca. 20% (Table1). Leaf mass per area was not different between leaf types. No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E, N~area~ or LMA either within or across leaf types.

## Leaf photosynthesis rates and photosynthetic capacity
Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P < 0.001, Table 2). Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light. Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure `r get1fig`a).

\
The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (P < 0.001 & P = 0.02, respectively), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). No effect of the warming treatment was detected on either parameter (Table 1). Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).  Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).  As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.  When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)

## Leaf water use efficiency
Leaf ITE was significantly greater in sun leaves than in shade leaves (P = 0.001, Table 2). Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).  ITE in sun leaves was reduced in the warming treatment (P = 0.021) but no effect was noted in shade leaves with low or high light. For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).  The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.  Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca. 2&permil; (p < 0.001, Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.

*leaf 13C with leaf N and what this stands for*

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a). In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves. 
\
The observed carbob isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7%pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments. Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P < 0.001, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature.

>>>>

##Leaf gas exchange, online carbon isotope discrimination and mesophyll conductance
Leaf gas exchange measurements were performed monthly through to the final harvest.  Measurements were taken on a representative sun and shade leaf for each tree during each measurement campaign.  The newest fully expanded leaf from the stem apex was chosen for gas exchange measurements and sun leaves were measured in the upper third of the canopy. In order to minimize any confounding effects of leaf age, shade leaves were measured first in the lower canopy then gradually higher during each campaign.  As shade leaves most likely developed slower this assured that older leaves in the lower canopy were avoided. The nearest leaf on each branch was sampled for measurement of predawn leaf water potential.
\
Prior to gas exchange measurements photosynthetic photon flux density (PPFD) was recorded at the individual leaf level and at the canopy position for each selected leaf.  A hand-held quantum sensor (LI-COR, Lincoln, NE, USA) was used to record leaf level PPFD to ensure that chosen leaves were positioned in the desired light environment, either full sun or full shade. A ceptometer (AccuPAR LP-80, Decagon Devices, Pullman, USA) was then used to measure 1 m integrated PPFD at the canopy height of each chosen leaf type. Five ceptometer readings were recorded within the canopy at random locations of the height of each selected leaf.  The mean of these readings was assumed to represent the local light environment of full sun and shade leaves for each tree. All measurements of PPFD and gas exchange were performed on full sun days between 10:00-14:30 h.

@@

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite. Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al. [-@evans1986carbon]:

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The ternary effect corrections (t) are desribed by:

@@

where E denotes the transpiration rate and $\text{g}^t_{ac}$ is the total conductance to CO~2~ diffusion to both the boundary layer and stomatal conductance.

@@

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R with WTC as a random effect. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models of A as a function of different physiological parameters were generated using bootstrapping methods with 999 simulations in the 'lme4' package [@bates2013lme4] in R. For non-linear mixed models models confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package in R.  Results were considered significant at P &leq; 0.05.

@@

## Leaf resource distribution
Across all measurement campaigns 1 m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure `r parfig`). The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).  This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types. Leaf N, on an area basis, was approximately 20% higher in sun leaves than in shade leaves (Table1). Leaf mass per area was not different between leaf types. No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E, N~area~ or LMA either within or across leaf types.

## Leaf photosynthesis rates and photosynthetic capacity
Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (Table 2). Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

\
The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). No effect of the warming treatment was detected on either parameter (Table 1). Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).  Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).  As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.  When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)

## Leaf water use efficiency
Leaf ITE was significantly greater in sun leaves than in shade leaves (P = 0.001, Table 2). Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).  ITE in sun leaves was reduced in the warming treatment (P = 0.021) but no effect was noted in shade leaves with low or high light. For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).  The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

\
Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca. 2&permil; (Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a). In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.

\
The observed carbon isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.

\
Mean g~m~ was higher in sun compared to shade leaves (+27%), under their local light environment ( Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature was detected.
